---
title: "Global Energy Transition & Market Reliability"
subtitle: "A cross-country comparison: United States · United Kingdom · Australia"
author: "Your Name"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    code-fold: true
    self-contained: true
  pdf:
    toc: true
    geometry: margin=2.5cm
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(plotly)
library(here)

combined <- read_csv(here("data/processed/combined_energy.csv"))
```

## Introduction

This report analyses energy generation trends across the United States, United Kingdom,
and Australia. The three markets represent distinct stages of the global energy transition,
from Australia's coal-heavy legacy grid to the UK's rapidly decarbonising mix.

Data spans 2019 to present, with Australian historical data available from 1999.

---

## Generation mix

```{r generation-mix}
#| fig-cap: "Monthly electricity generation by source and country (GWh)"

combined |>
  group_by(country, date, source) |>
  summarise(gwh = sum(generation_gwh), .groups = "drop") |>
  ggplot(aes(x = date, y = gwh, fill = source)) +
  geom_area(position = "stack", alpha = 0.85) +
  facet_wrap(~country, ncol = 1, scales = "free_y") +
  scale_fill_brewer(palette = "Set2") +
  labs(x = NULL, y = "GWh", fill = "Source") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

---

## Renewable penetration

```{r renewables}
#| fig-cap: "Renewable share of total generation by country (%)"

combined |>
  group_by(country, date) |>
  summarise(
    pct = sum(generation_gwh[renewable]) / sum(generation_gwh) * 100,
    .groups = "drop"
  ) |>
  ggplot(aes(x = date, y = pct, colour = country)) +
  geom_line(linewidth = 0.9) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_colour_manual(values = c("#2c7bb6", "#d7191c", "#1a9641")) +
  labs(x = NULL, y = "Renewable share (%)", colour = "Country") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
```

---

## Cross-country comparison

```{r comparison}
#| fig-cap: "Total monthly generation by country (GWh)"

p <- combined |>
  group_by(country, date) |>
  summarise(total_gwh = sum(generation_gwh), .groups = "drop") |>
  plot_ly(x = ~date, y = ~total_gwh, color = ~country, type = "scatter",
          mode = "lines", colors = c("#2c7bb6", "#d7191c", "#1a9641")) |>
  layout(
    xaxis = list(title = ""),
    yaxis = list(title = "GWh"),
    legend = list(orientation = "h")
  )
p
```

---

## Emissions intensity

*Coming soon — requires merging Australian emissions data with UK/US equivalents.*

---

## Conclusions

*To be completed after analysis.*

---

## Data sources

- **UK**: GridWatch — [gridwatch.templar.co.uk](https://gridwatch.templar.co.uk)
- **US**: EIA Hourly Electric Grid Monitor — [eia.gov](https://www.eia.gov/electricity/gridmonitor/)
- **AU**: AEMO All Regions — [aemo.com.au](https://www.aemo.com.au)
